\documentclass[12pt]{article}

\input{memo_setup}
\input{abbrev}
\usepackage{ewv_environ}      %   %% Define some new environements                            

%\graphicspath{{sigmac_0999/}}

\begin{document}                                                 

\title{BTeV Document Database Design and Interface}

\author{
        Eric W. Vaandering\\  
        Vanderbilt University\\
        \\
        {\small\url{http://www.hep.vanderbilt.edu/~somewhere}}
       }
        

\date{27 November 2001 \\
      Revised: 11 December 2001}

\maketitle

\begin{abstract} 
In this memo, we describe the design and implementation of the BTeV document
database.
\end{abstract}

\section{Design considerations}

We needed a document database that satisfies a number of requirements:
\begin{itemize}
\item{Each document can have multiple revisions, and old versions are still
available}
\item{Each revision of each document can have multiple files. This accomodates
multiple file types (source and  presentable) and/or child files (especially
useful for trees of HTML  documents)}
\item{All documents are kept on the BTeV web server, not local copies 
so we don't suffer from link rot}
\item{The ability to upload documents by browser upload or forcing the 
document database to download via http.}
\item{The ability of a single database to store talks from group meetings,
conference talks and proceedings, and publications \emph{and} to present these
special cases in a way that makes sense for that special case.}
\end{itemize}

There are several security issues we want to resolve too:
\begin{itemize}
\item{The ability to not only have public/private documents but also 
documents that are accessible to subgroups (like RTES) which don't know 
the BTeV password. We also want to have documents that are only 
accessible to sub-groups (like the executive council).}
\item{Users should only know that a document exists if they have 
privileges to view it.}
\item{The ability to easily move documents, or just certain versions of 
documents, from restricted to public and vice-versa.}
\end{itemize}

\section{Interface}

The document database is completely accessible via the web. HTML and HTML forms
provide the user interface and CGI scripts 
handle all accesses to the underlying database. 

While a search engine is not yet included in the product, one is planned.

These scripts are all written in Perl using the Perl::CGI package and 
the Perl::DBI database access modules. There is no reason 
additional access programs can't be written using PHP, C++, or any other 
language. 

\section{Database}

The database design is essentialy finished, although small additions will
probably be needed for conference presentations. (See \secref{extend}).

The document 
database is implemented as a number of different tables. Because each 
document can have multiple revisions and each revision can have 
multiple files (and authors and topics), there are three main tables 
Document, DocumentRevision, and DocumentFile. The hierarchy of these 
tables, their elements, and the auxiliary tables are shown in figref.

The database design we are using is in the ``third normal form.'' For a longer
discussion of what this means and how a database is properly designed
see~\cite{YargerReeseKing:1999}. Essentially it boils down to three requirements:
\begin{enumerate}
\item{Each field of a database contains only one piece of information}
\item{Any field which has values in common across multiple entries should use a
unique identifier and another table}
\item{No two fields in a table should depend on eachother}
\end{enumerate}
To take a simple example, the field containing the document type should contain
a unique identifier, not the name of the type because the name might have been
mis-spelled (rule 2). The institution of the person who wrote the document
something is not a property of the document, but of the person (rule 3).

Fields shown in red are indexed for fast look up. Other fields must be searched
on which is more time consuming. Each index takes up some  amount of disk
space, so the number of indices should be kept to a minimum. The ``tree''-like
symbols indicate the one-to-many relationships within the database. For
instance, there can (and will) be many revisions of a  single document.  Rule 1
mandates that there can be no ``trees'' with branches on both ends. The ID
numbers in each entry indicate what other entries they are associated with. In
addition, each table has a timestamp which is not shown.

This database is implemented using MySQL; the server is running on
\url{fnsimu1.fnal.gov}.  

\section{File system and security}

All documents are placed into a single directory tree. There is no 
division based on security or topic since these are fluid designations 
(and each document can have multiple values for these settings). The 
file system is arranged only by document number. Each revision of each 
document is placed in its own directory, which means that most 
directories will only have a small number of files. However, the 
advantage of this scheme is two-fold. First, each of the files of a 
particular revision need not be renamed. Second, access to each 
revision can be controlled with a potentially unique .htaccess file.

To avoid possible issues with large numbers of files in a single 
directory, the document directories are further divided into groups of 
100. For example, imagine that the 3rd revision of the 1029th document 
contains one file, text.ps. Then the file tree looks like this:

\texttt{\$DOCROOT/0010/001029/003/text.ps}

As you can see, the file system is designed for up to 999,999 documents 
each of which can have up to 999 versions.

\section{Planned Extensions}\seclabel{extend}

\end{document}




